---
title: "Red Wine"
author: "Melinda K. Higgins, PhD."
date: "February 16, 2017"
output:
  pdf_document: default
  html_document: default
  github_document: default
  word_document: default
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# define libraries
library(ggplot2)
```

## Read in data

NOTE: There are 2 formatting issues with the data as downloaded from the UCI data repository. Even though the filename extension is CSV (usually "comma-delimited" format), 

1. the variable names have multiple words separated by spaces which many/most import routines will have problems with and
2. the "delimiter" or symbol used to separate data fields (aka, the columns) is the semicolon `;` so we need to specify this.

If it was a simple comma-delimtied file the function `read.csv()` would work fine. However, to address the issue of the semicolon, we need to use `read.table()` and provide the delimiter specifically.

NOTE: `read.table()` also "fixes" the variable names such that the spaces are filled in using a period `.`. For example, the 1st variable "fixed acidity" is converted to `fixed.acidity` in the final dataframe.

```{r}
redWine <- read.table("winequality-red.csv",
                      header=TRUE,
                      sep=";")
```

## Run a summary of the Red Wine Data

```{r}
summary(redWine)
```

## Histogram of Fixed Acidity using `ggplot()`

```{r}
ggplot(aes(fixed.acidity), data=redWine) +
  geom_histogram()
```

## Density Overlay with Histogram & add color

```{r}
ggplot(aes(fixed.acidity), data=redWine) + 
  geom_histogram(aes(y =..density..), 
                 breaks=seq(4, 16, by = 0.5), 
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  geom_density(col=2) + 
  labs(title="Histogram for Fixed Acidity") +
  labs(x="Fixed Acidity", y="Proportion")
```

