---
title: "Red Wine"
author: "Melinda K. Higgins, PhD."
date: "February 16, 2017"
output:
  word_document: default
  html_document: default
  pdf_document: default
  github_document: default
---

```{r setup}
# code chunk setting for RMD file
knitr::opts_chunk$set(echo = TRUE)

# load packages for RMD file
# any functions/code you use in your
# RMD file must have the correct package library
# loaded BEFORE you try to use the command/function
# or you will get errors when you knit the RMD file
library(ggplot2)
```

## Read in data

NOTE: There are 2 formatting issues with the data as downloaded from the UCI data repository. Even though the filename extension is CSV (usually "comma-delimited" format), 

1. the variable names have multiple words separated by spaces which many/most import routines will have problems with and
2. the "delimiter" or symbol used to separate data fields (aka, the columns) is the semicolon `;` so we need to specify this.

If it was a simple comma-delimtied file the function `read.csv()` would work fine. However, to address the issue of the semicolon, we need to use `read.table()` and provide the delimiter specifically.

NOTE: `read.table()` also "fixes" the variable names such that the spaces are filled in using a period `.`. For example, the 1st variable "fixed acidity" is converted to `fixed.acidity` in the final dataframe.

```{r}
# make sure the code to read in the
# data is in your RMD file otherwise
# when you knit, you will get errors
redWine <- read.table("winequality-red.csv",
                      header=TRUE,
                      sep=";")
```

## Run a summary of the Red Wine Data

```{r}
summary(redWine)
```

## Histogram of Fixed Acidity using `ggplot()`

```{r}
# aes() defines which variable you want the 
# histogram for.
ggplot(aes(fixed.acidity), data=redWine) +
  geom_histogram()
```

## Density Overlay with Histogram & add color

The code below adn be modified as you wish to make more interesting histograms with an overlaid density curve.

```{r}
# remember aes(..density..) is what you use to
# change from a frequency based (counts) histogram
# to one based on the proportions or "density"
# breaks=seq(4, 16, by = 0.5) defines the breakpoints
# we want to see for this variable "fixed.acidity"
# col="red" defines the outline color of each bar
# fill="green" defines the color inside each bar
# alpha=.2 is the transparency level for the fill color
# the overlaid geom_density() col is also "red" (color 2)
# finally custom title, xlabel and ylabel are provided
ggplot(aes(fixed.acidity), data=redWine) + 
  geom_histogram(aes(y =..density..), 
                 breaks=seq(4, 16, by = 0.5), 
                 col="red", 
                 fill="green", 
                 alpha = .2) + 
  geom_density(col=2) + 
  labs(title="Histogram for Fixed Acidity") +
  labs(x="Fixed Acidity", y="Proportion")
```

## Example Scatterplot Matrix

```{r}
# load car package - to use scatterplotMatrix()
library(car)

# make scatterplot matrix for quality and 
# some of the first variables
scatterplotMatrix(~ quality + fixed.acidity +
                     volatile.acidity + citric.acid +
                     residual.sugar, data=redWine)
```

## try a linear model and summarize

```{r}
model1 <- lm(quality ~ fixed.acidity +
                     volatile.acidity + citric.acid +
                     residual.sugar, data=redWine)
summary(model1)
```

## run a 2nd model and compare

Drop all non-significant variables - so keep "volatile.acidity" and run again

```{r}
model2 <- lm(quality ~ volatile.acidity, data=redWine)
summary(model2)
```

## Compare these 2 linear models

We can use the `stargazer` package to compare 2 similar models. In order for the tables to work correctly for HTML output, you need to do 2 things: (1) add `results="asis"` to the `R` code chunk option and (2) add `type="html"` as an option in the `stargazer()` function call.

_**WARNING**: This approach will NOT work if you "knit" to DOCX or PDF. It only works for HTML output._

```{r, results="asis"}
library(stargazer)
stargazer(model1, model2, 
          title="Comparison of 2 Regression outputs",
          type="html")
```

If you want a format that will work for all of the output formats (HTML, DOCX and PDF), use `type="text"` option which will give simple R text-based output.

```{r}
stargazer(model1, model2, 
          title="Comparison of 2 Regression outputs",
          type="text")
```

## look at some diagnostic plots

```{r}
residualPlots(model1)
avPlots(model1, id.n=3, id.cex=0.7)
qqPlot(model1, id.n=3)
influenceIndexPlot(model1, id.n=3)
```

## Other diagnostics

```{r}
outlierTest(model1)
ncvTest(model1)
vif(model1)
```

